---
title: "Math 615 homepage"
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
wk1.date <- as.Date("2022-08-23")
```

This is the course website for <span style="color:#6f1d1b">**MATH 615 - Statistical Methods for Graduate Research**</span> course taught by [Dr. Robin Donatello](www.norcalbiostat.com) in Fall 2022 at California State University, Chico. This landing page is used for posting of regular announcements and information for students of the class.

# Revisions for Fall 22 in progress


----

_Last Updated: `r format(Sys.time(), "%a %b %d %X")`_

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr);library(readxl); library(kableExtra)
options(knitr.table.format = "html", knitr.kable.NA = '') 
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE) 
```

# Schedule  {.tabset .tabset-fade .tabset-pills}

## Weekly Overview 

```{r}
# Read in schedule
sched <- read_excel("schedule_615.xlsx", sheet="weekly_schedule") %>% 
  mutate(Date = format(date, "%m/%d")) %>% 
  select(Week, order, Date, Topic, Assignment)

# set colors for schedule
tn <- sched %>% filter(!is.na(Week)) %>% 
  select(Week) %>% 
  mutate(start=0, end=0)

for(t in 1:16){
  tn$start[t] <- min(which(floor(sched$order)==t))
  tn$end[t]   <- max(which(floor(sched$order)==t))
}


wk.col <-  "background-color: #654378; color: #fff;"
hlt.col <-  "background-color: #50a18d; color: #fff;"

```

```{r}
sched %>% select(-order) %>%
  kbl(align="lclc", escape=FALSE) %>%
  kable_paper("hover", full_width=F, html_font = "Ariel",
              position='center', font_size=14) %>%
  column_spec(1, width="5%") %>% 
  column_spec(2, width="5%", border_right = TRUE) %>% 
  column_spec(3, width="20%") %>% 
  column_spec(4, width="20%") %>% 
  pack_rows("Week 1", tn$start[1],   tn$end[1],  label_row_css = hlt.col) %>%
  pack_rows("Week 2", tn$start[2],   tn$end[2],  label_row_css = wk.col) %>%
  pack_rows("Week 3", tn$start[3],   tn$end[3],  label_row_css = wk.col) %>%
  pack_rows("Week 4", tn$start[4],   tn$end[4],  label_row_css = wk.col) %>%
  pack_rows("Week 5", tn$start[5],   tn$end[5],  label_row_css = wk.col) %>%
  pack_rows("Week 6", tn$start[6],   tn$end[6],  label_row_css = wk.col) %>%
  pack_rows("Week 7", tn$start[7],   tn$end[7],  label_row_css = wk.col) %>%
  pack_rows("Week 8", tn$start[8],   tn$end[8],  label_row_css = wk.col) %>%
  pack_rows("Week 9", tn$start[9],   tn$end[9],  label_row_css = wk.col) %>%
  pack_rows("Week 10", tn$start[10],   tn$end[10],  label_row_css = wk.col) %>%
  pack_rows("Week 11", tn$start[11],   tn$end[11],  label_row_css = wk.col) %>%
  pack_rows("Week 12", tn$start[12],   tn$end[12],  label_row_css = wk.col) %>%
  pack_rows("Week 13", tn$start[13],   tn$end[13],  label_row_css = wk.col) %>%
  pack_rows("Week 14", tn$start[14],   tn$end[14],  label_row_css = wk.col) %>%
  pack_rows("Week 15", tn$start[15],   tn$end[15],  label_row_css = wk.col) %>%
  pack_rows("Finals",  tn$start[16],   tn$end[16],  label_row_css = wk.col)
```



## Google Calendar (Due Dates)

For those of you that like the calendar view of your schedule. Due to schedule adjustments during the semester these dates may not match exactly with the schedule. In those cases please bring any descrepencies to my attention via Discord. 

> `r emo::ji("idea")` Subscribe to this calendar by clicking the plus button in the bottom right corner!

<iframe src="https://calendar.google.com/calendar/embed?src=mail.csuchico.edu_bpvuoivh18qp7ui1263m0dju3o%40group.calendar.google.com&ctz=America%2FLos_Angeles" style="border-width:0" width="1000" height="800" frameborder="0" scrolling="no"></iframe>

