---
title: 'Math 615 Materials'
output:
  html_document:
    toc: yes
    toc_float: yes
---
<br>
```{r, message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(stringr); library(dplyr); library(readxl)
library(knitr); library(kableExtra); library(tidyr)
options(knitr.kable.NA = '')

lec  <- data.frame(filenames = paste0('lecture/', list.files("lecture", pattern='*.pdf|*.html|*.docx')))
read <- data.frame(filenames = paste0('reading/', list.files("reading", pattern='*.pdf|*.html|*.docx')))
hw   <- data.frame(filenames = paste0('hw/', list.files("hw", pattern='template|*.pdf|*.html|*.docx')))

out.lec <- lec %>% mutate(name = gsub("lecture/",  "", filenames), 
                          type = str_extract(lec$filenames,"\\.[a-z]{3}"),
                          type = ifelse(type==".htm", "html", type),
                          type = ifelse(type=="doc", "docx", type),
                          type = ifelse(type==".pdf", "pdf", type),
                          nam = gsub(".html|.pdf|.docx", "", name), 
                          link = sprintf("[%s](%s)",type, filenames))

out.read <- read %>% mutate(name = gsub("reading/",  "", filenames), 
                            name = gsub(".pdf", "", name),
                            link = sprintf("[%s](%s)",name, filenames))

out.hw <- hw %>% mutate(name = gsub("hw/|_template",  "", filenames), 
                          type = str_extract(hw$filenames,".{3}$"),
                          type = ifelse(type=="tml", "html", type),
                          type = ifelse(type=="ocx", "docx", type),
                          nam = gsub(".html|.pdf|.Rmd|.docx", "", name), 
                          link = sprintf("[%s](%s)",type, filenames))
```

This page contains links to all materials needed for this class. 

# Lecture Notes

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
x = out.lec %>% select(link, type, nam) %>% 
    reshape2::dcast(nam~type, value.var="link") %>% 
    kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", "bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```

# Collaborative Notes
Links are to notes on HackMD

```{r}
x <- read_excel("schedule_615.xlsx", sheet="weekly_schedule") %>% 
  select(`Collab Notes`) %>% 
  mutate(Item = gsub("\r\n", "@", `Collab Notes`)) %>% 
  select(-`Collab Notes`) %>%
  na.omit() %>%
  separate_rows(Item, sep = "@") %>% 
  kbl(align="ll", escape=FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive" ,"bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```

# Programming

* [Collaborative R notes on Hack MD](https://hackmd.io/@norcalbiostat/r_notes)
* [Collaborative SPSS notes on Hack MD](https://hackmd.io/@norcalbiostat/spss_notes)
      

# Homework files

To download RMD template files, right click and "save as". Do not open in a browser then save. 
Word doc templates are built off the same file as the R template, so it may contain irrelevant verbiage that you can delete. 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
x = out.hw %>% select(link, type, nam) %>% 
    reshape2::dcast(nam~type, value.var="link") %>% 
    kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive" ,"bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```


# Articles & Readings

Some of these are required readings, some are for your information and curiosity. 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
x = out.read %>% select(link) %>% kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive" ,"bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```

Add non PMA6 readings from schedule. 

# Statistics
* [OpenIntro Statistics](https://www.openintro.org/book/os/) Free online learning material for intro statistics. 
* [Applied Statistics Notebook](https://norcalbiostat.github.io/AppliedStatistics_notes/index.html) (Abbreviated as **AS Notebook**) 

