---
css: ../css/customh5.css
output:
  html_document:
    highlight: tango
    theme: flatly
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(stringr); library(dplyr); library(reshape2)
library(knitr); library(kableExtra)

lec <- data.frame(filenames = paste0('lecture/', list.files("lecture", pattern='*.pdf|*.html')))
read <- data.frame(filenames = paste0('reading/', list.files("reading", pattern='*.pdf|*.html')))
hw <- data.frame(filenames = paste0('hw/', list.files("hw", pattern='*.pdf|*.html')))

out.lec <- lec %>% mutate(name = gsub("lecture/",  "", filenames), 
                          type = str_extract(lec$filenames,"\\.[a-z]{3}"),
                          nam = gsub(".html|.pdf", "", name), 
                          link = sprintf("[%s](%s)",type, filenames))

out.read <- read %>% mutate(name = gsub("reading/",  "", filenames), 
                            link = sprintf("[%s](%s)",name, filenames))

out.hw <- hw %>% mutate(name = gsub("hw/",  "", filenames), 
                          type = str_extract(hw$filenames,"\\.[a-z]{3}"),
                          nam = gsub(".html|.pdf", "", name), 
                          link = sprintf("[%s](%s)",type, filenames))
```

<br>

## Textbook based lecture notes 

* The [Applied Statistics Notebook](https://norcalbiostat.github.io/AppliedStatistics_notes/index.html) (Abbreviated as **AS Notebook**) is a general set of notes that span multiple topics from 315, 615 and 456.  
* If you are rusty on your basic statistics (i.e. what is a normal distribution, how do you interpret variance) then look into the [Open Intro Statistics](https://www.openintro.org/stat/textbook.php?stat_book=os) textbook. The PDF version is free and has links to videos in it etc. 

## One-off lecture notes 

* Other sets of lecture notes are listed below. 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
x = out.lec %>% select(link, type, nam) %>% dcast(nam~type, value.var="link") %>% 
    kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", "bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```

## Reading Materials
```{r, message=FALSE, warning=FALSE, echo=FALSE}
x = out.read %>% select(link) %>% kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive" ,"bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```

## Homework Assignment Files
```{r, message=FALSE, warning=FALSE, echo=FALSE}
x = out.hw %>% select(link, type, nam) %>% dcast(nam~type, value.var="link") %>% 
    kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive" ,"bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```


