---
output:
  html_document:
    toc: yes
    toc_float: yes
---
<br>
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(stringr); library(dplyr); library(reshape2)
library(knitr); library(kableExtra)
options(knitr.kable.NA = '')

lec <- data.frame(filenames = paste0('lecture/', list.files("lecture", pattern='*.pdf|*.html|*.docx')))
read <- data.frame(filenames = paste0('reading/', list.files("reading", pattern='*.pdf|*.html|*.docx')))
hw <- data.frame(filenames = paste0('hw/', list.files("hw", pattern='template|*.pdf|*.html|*.docx')))

out.lec <- lec %>% mutate(name = gsub("lecture/",  "", filenames), 
                          type = str_extract(lec$filenames,"\\.[a-z]{3}"),
                          type = ifelse(type=="tml", "html", type),
                          type = ifelse(type=="doc", "docx", type),
                          nam = gsub(".html|.pdf|.docx", "", name), 
                          link = sprintf("[%s](%s)",type, filenames))

out.read <- read %>% mutate(name = gsub("reading/",  "", filenames), 
                            name = gsub(".pdf", "", name),
                            link = sprintf("[%s](%s)",name, filenames))

out.hw <- hw %>% mutate(name = gsub("hw/|_template",  "", filenames), 
                          type = str_extract(hw$filenames,".{3}$"),
                          type = ifelse(type=="tml", "html", type),
                          type = ifelse(type=="ocx", "docx", type),
                          nam = gsub(".html|.pdf|.Rmd|.docx", "", name), 
                          link = sprintf("[%s](%s)",type, filenames))
```

# Notes
We will be using a combination of notes from the list below


```{r, eval=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
x = out.lec %>% select(link, type, nam) %>% dcast(nam~type, value.var="link") %>% 
    kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", "bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```

* Dr. D's [Math 315 course notes](https://norcalbiostat.github.io/MATH315/reading/RAD_course_notes_f19.pdf) 
* [Applied Statistics Notebook](https://norcalbiostat.github.io/AppliedStatistics_notes/index.html) (Abbreviated as **AS Notebook**) 
* Lec00: Intro to the class [[HTML]](lecture/lec00_intro_class.html)[[PDF]](lecture/lec00_intro_class.pdf)
* [[Lec01]](https://hackmd.io/@norcalbiostat/math615_lec01): Data architecture
* [[Lec02]](https://hackmd.io/@norcalbiostat/math615_lec02): Formulating Research Questions
* [[Lec03]](https://hackmd.io/@norcalbiostat/math615_lec03): Preparing Data for Analysis
* [[Lec04]](https://hackmd.io/@norcalbiostat/math615_lec04): Describing Distributions and Relationships
* [[Lec05]](https://hackmd.io/@norcalbiostat/math615_lec05): Probability and Sampling Distributions 
* [[Lec06]](https://hackmd.io/@norcalbiostat/math615_lec06): Interval Estimation and Hypothesis Testing 


----

# Homework files

To download RMD template files, right click and "save as". Do not open in a browser then save. 
Word doc templates are built off the same file as the R template, so it may contain irrelevant verbiage that you can delete. 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
x = out.hw %>% select(link, type, nam) %>% dcast(nam~type, value.var="link") %>% 
    kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive" ,"bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```


# Articles
Some of these are required readings, some are for your information and curiosity. 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
x = out.read %>% select(link) %>% kable(format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive" ,"bordered"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```
