---
title: 'Project Information'
date: "Last Updated `r Sys.time()`"
output:
  html_document:
    highlight: tango
    theme: flatly
    toc: yes
    toc_float: yes
css: css/customh5.css
---

# Overview

Throughout the entire semester you will be analyzing a data set with the end goal of creating a poster presentation to present your research. Often your homework serves as a first draft of exploration into your research questions, and helps you build your story. Homework assignments alone will not be sufficient for a good analysis. 

# Poster preparation

You will organize your research in stages by building a "powerpoint" type presentation. This will help you concisely explain your research topic and findings in a way that you can more easily translate (fit) onto a poster. 

### Instructions
* Create a Google Slides presentation directly into the `Poster prep (slides)` folder in our Google Drive. 
  - Name your file using `first_last` name only. 
* The required contents for each slide are listed below under **Stages**

### Stages
* At each stage you should update anything you need to change, such as correctly making graphics, writing results correctly etc. 
* This includes reviewing the current set of feedback and make the appropriate corrections. 
    - Once you have addressed a particular piece of feedback mark it as "resolved" in Google Slides. 

### List of slide contents
Your presentation should be limited to approximately 17 slides. The contents of each slide are listed below. 


#### <span style="color:blue">**Stage 1: Background and sample characteristics**</span>
Lecture notes on writing about empirical research: https://norcalbiostat.github.io/MATH615/lecture/lec02b_writing_em.html 

1. Title 
2. Introduction 
3. Background & lit review. 
4. Explain the research problem / topic area
5. State your Research Question as a testable hypothesis. 
6. Introduce the data. Where does it come from, how was it collected. How many records? 
7. Description of measures (variables) being used. Explain any recoding and/or altering that was done. 
    - This refers to significant changes to the data (making new variables, dichotomizing or categorizing existing variables)
    - Do NOT discuss setting missing codes such as 98 and 99 to `NA`. 
    - Do NOT talk about applying labels
    - Do NOT use variable names. 
    - Restrict this list to only the variables you are using in your analysis. This is likely not to be more than 4-6. 
8. Fully describe your main explanatory variable univariately using summary statistics and graphics. 
9. Fully describe your main response variable univariately using summary statistics and graphics. 

#### <span style="color:blue">**Stage 2: Bivariate analysis**</span>

Two analyses, one per slide. Each slide should contain the following:  

1. A fully labeled and cleaned up plot demonstrating the bivariate relationship under consideration.
2. The results of an appropriate bivariate analysis. 
    - This should be a conclusion written in English, includes a confidence interval and a p-value. 
    - You also need to state the name of the test you used. 

#### <span style="color:blue">**Stage 3: Multivariable analysis**</span>

You are trying to understand the relationship between your explanatory and response variable, in presence of information contained in other variables. 

**Slide 12: Testing for moderators**

* Moderation:
    - If you found a moderator: Remove the bivariate graph and write up from your poster and replace with explanation of moderator and graph of significant bivariate at the 3rd variable level.
    - If you did NOT find a moderator: Include a slide stating what relationships you tested for moderation and with what variable to see if it was a moderator, explaining that it was not a moderator. No graphs needed

<!---
* Confounding: 
    - If you found a confounder: on slide with bivariate relationship that was confounded put a bullet point that explains what variable confounded the relationship and put the written results that are shown in the confounding assignment on how to write results. Make sure to explain if it helped improve prediction of response variable. 
    - If you did NOT find a confounder: Include a slide stating what relationships you tested for confounds and with what variable to see if it was a confound, explaining that it was not a moderator. Make sure to explain if it helped improve prediction of response variable. 
--->

**Slide 13: Results from a Multivariable Model**

* Build a multivariable model by adding additional predictors to the model. 
    - See the lecture notes on model building as guidance. 
    - Include any variables that were found to be significantly associated with the outcome
    - If you found a moderator, your model should include an interaction term with your moderating variable.
* A table or a graphic of the regression coefficients (or Odds Ratios) must be presented. 
    - Make tables in Excel or Powerpoint directly. Do not copy R output
* At least one coefficient, the primary explanatory variable, must be interpreted in context of the problem. 


**Slide 14. Discussion**

* Here you will explain what your graphical and inferential results tell you about your topic.
* Discuss if your research hypothesis was supported, if it was not, why you think that might be
* Explain the overall story/trend/what you learned when you consider your univariate, bivariate & multivariate results about your topic.
* Compare your results to previous research results. Do they agree or disagree? 

**Slide 15. Implications**

* What are the practical implications of your results?
* What could others do with your findings? 
* What future research needs to be conducted?
    - This needs to be more specific than "other variables could be explored". Which variables and why? What other research articles indicate that those other variables are relevant? 

**Slide 16. Limitations**

* Who are the results of this study generalizable to? (i.e. a subset of individuals?)
* Were there any model assumptions that were not upheld? 
* If this is an observational study, you should make a statement about the findings are associations and not causal in nature
* Are there other factors that could exlain your response variable that you did not include in your model? 


**Slide 17. References**  

* You can use smaller font to get all references on one slide.
* Use references from research plan, and any additional references gathered along the way. 
* Make sure these are correctly done in APA format.
* Proper citations for 
    - [Add Health](http://www.cpc.unc.edu/projects/addhealth/faqs/addhealth#how-do-i-cite-the-add-health-contractual-data-)
    - *R*: Type `citation()`
    - [R Studio](https://support.rstudio.com/hc/en-us/articles/206212048-Citing-RStudio)
    - [SPSS](http://www-01.ibm.com/support/docview.wss?uid=swg21476197)
    - [How to cite software in Text](http://blog.apastyle.org/apastyle/2015/01/how-to-cite-software-in-apa-style.html)
    


----

# Posters Presentations

* You will transfer all findings into a research poster, print the poster, and then present your research to your classmates. 
    - We will split the class in half. When not presenting, you will be scoring the other half of the class's posters. 
* All posters must be built in Powerpoint. No exceptions.
* Use these [poster guidelines](https://norcalbiostat.netlify.com/lec/poster_guidelines/) to see templates and guidance on how to make an effective poster.


### Draft version
* This draft is graded on how complete the poster is. You should consider this a draft that you would circulate to your colleagues for final review and comments. 
* Use the [evaluation criteria](https://norcalbiostat.netlify.com/lec/poster_guidelines/#evaluation-criteria) as a guide for what I am looking for.

**Submission and Peer Review Instructions**

* Upload your draft (PPTX or PDF) to the `Google Drive/Project Poster/Draft` folder
    - Name this file `authorname_poster_draft.pptx`
* To conduct your peer review: 
    - Download your Assigned posters to review. 
    - Use the comment  feature to write the comments directly into the poster file itself. 
    - provide at least two specific comments for each
        - something that can be improved
        - something that is good
    - Save this file with your initials at the end of the file name.  `authorname_poster_draft_RAD.pptx` 
* Upload your revision to the `Google Drive/Project Poster/Draft/Reviewed` folder

<!---
<div id="draft" class="collapse"> 
#### Poster drafts to review. 
```{r, message=FALSE, warning=FALSE}
library(stringr); library(dplyr)
library(knitr); library(kableExtra)

dta <- data.frame(filenames = paste0('PosterDraft/', list.files("PosterDraft/")))

out <- dta %>% mutate(name = gsub("PosterDraft/",  "", filenames),
                      link = sprintf("[%s](%s)",name, filenames)) %>%
               select(link)

x = kable(out, format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)
```

</div>

--->


### Final Version
* Upload your final poster (as printed) to Blackboard Learn
    - PDF preferred. PPTX ok.
* Scoring will be done via [Google Forms](https://goo.gl/forms/QORgO0jjqNzp3T5i2). It looks exactly like the [evaluation criteria](https://norcalbiostat.netlify.com/lec/poster_guidelines/#evaluation-criteria).
    - When not presenting, you will walk around and learn about others research. Ask them questions, fill out the evaluation on your phone as you go. Printed copies will be available upon request. 
* The final grade will be determined by
    - Base score = **30%** student score + **70%** instructor score **+ bonus pts** for best in show awards **- 1** point penalty for each day late. 

<!---

<button data-toggle="collapse" data-target="#final">Show/Hide Final Versions</button>
<div id="final" class="collapse">

```{r, message=FALSE, warning=FALSE, echo=FALSE}
dta <- data.frame(filenames = paste0('FinalPoster/', list.files("FinalPoster/")))

out <- dta %>% mutate(name = gsub("FinalPoster/",  "", filenames),
                      link = sprintf("[%s](%s)",name, filenames)) %>%
               select(link)

x = kable(out, format="html") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                  full_width=FALSE,  position = "left")
gsub("<thead>.*</thead>", "", x)

```

</div>
--->

