---
title: "Schedule"
output: html_document
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr);library(readxl); library(kableExtra);library(stringr); library(pander)
options(knitr.table.format = "html", knitr.kable.NA = '') 
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE) 
```

# Last Updated: `r format(Sys.time(), "%a %b %d %X")` {.tabset .tabset-fade .tabset-pills}

## Overview

<p>
Unless specified otherwise, all assignments are due at midnight on the due date listed.  

Don't forget to check the weekly details for information on how to prepare and links to other materials. 
</p>
```{r}
# Read in schedule
rawdat <- read_xlsx("schedule_615.xlsx", sheet='weekly_schedule')

# Set semester start date and format weeks nicely
start.date <- as.Date("2019-08-25")

# which rows have week numbers (manually add line number for break week and finals)
wn <- c(which(!is.na(rawdat$wknum)),41, 46) %>% sort() 

# generate list of weekdates
wkdt <- seq.Date(start.date, start.date+7*16, by='week')

# format for display
weekdates <- paste0("Week ", rawdat$wknum[wn], " (", format(wkdt, "%m/%d") , ")") 
weekdates[14] <- "Fall Break (11/24)"
weekdates[17] <- "Finals Week (12/15)"

# Filter rows to display on schedule
out <- rawdat[which(rawdat$Order==1.1):nrow(rawdat), ]%>% select(Topic:Assignments)

#week <- grep("midterm|final", out$Topic, ignore.case=TRUE)

# add bullets to materials
#mats <- which(!is.na(out$Materials))
#out$Materials[mats] <- paste0("<ul><li>", out$Materials[mats], "</li></ul>")
#out$Materials[mats] <- gsub("\r\n", "</li><li>", out$Materials[mats])

# add bold to due dates
out$Assignments <- gsub(" (", " (<b>", out$Assignments, fixed=TRUE)
out$Assignments <- gsub(" )", "</b>)", out$Assignments, fixed=TRUE)
out$Quiz <- gsub(" (", " (<b>", out$Quiz, fixed=TRUE)
out$Quiz <- gsub(" )", "</b>)", out$Quiz, fixed=TRUE)

wk.labs <-  "background-color: #4d6600; color: #fff;"
hlt.wks <-  "background-color: #754b13; color: #fff;"
```

```{r, results='asis'}
out %>% 
  kable(align="lccclc", escape=FALSE) %>% 
  kable_styling(bootstrap_options = c("hover", "striped", "responsive", "bordered")) %>% 
  pack_rows(weekdates[1], wn[1],  wn[2]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[2], wn[2],  wn[3]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[3], wn[3],  wn[4]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[4], wn[4],  wn[5]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[5], wn[5],  wn[6]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[6], wn[6],  wn[7]-1, label_row_css =hlt.wks) %>%
  pack_rows(weekdates[7], wn[7],  wn[8]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[8], wn[8],  wn[9]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[9], wn[9],  wn[10]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[10], wn[10],  wn[11]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[11], wn[11],  wn[12]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[12], wn[12],  wn[13]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[13], wn[13],  wn[14]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[14], wn[14],  wn[15]-1, label_row_css =hlt.wks) %>%
  pack_rows(weekdates[15], wn[15],  wn[16]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[16], wn[16],  wn[17]-1, label_row_css =wk.labs) %>%
  pack_rows(weekdates[17], wn[17],  wn[17], label_row_css =hlt.wks) %>%
  column_spec(1, width = "35%") %>% 
  column_spec(2, width = "6%") %>%
  column_spec(3, width = "6%") %>%
  column_spec(4, width = "18%") %>% 
  column_spec(5, width = "35%")
```


## Calendar

For those of you that like the calendar view of your schedule. Due to schedule adjustments during
the semester these dates may not match exactly with the schedule. In those cases please bring any 
descrepencies to my attention via Slack. 

<iframe src="https://calendar.google.com/calendar/embed?src=mail.csuchico.edu_bpvuoivh18qp7ui1263m0dju3o%40group.calendar.google.com&ctz=America%2FLos_Angeles" style="border-width:0" width="1000" height="800" frameborder="0" scrolling="no"></iframe>

