---
title: "Week 01: Introduction to the Class"
output:   
  html_document: 
    toc: yes
    toc_float: true
    toc_depth: 2
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr);library(readxl); library(kableExtra); library(tidyr)
library(stringr)
options(knitr.table.format = "html", knitr.kable.NA = '') 
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE) 
```


----

# Learning Path

## Where we've been
Summer Break! 

## Where we're at
Ready to learn


## Objectives for the week

* Get connected to all the class materials & tools
* Learning how to format data in spreadsheets
* Choose a research project data set and review its codebook

---

```{r}
read_excel("schedule_615.xlsx", sheet="weekly_schedule") %>% 
  mutate(Date = format(date, "%m/%d")) %>% 
  filter(floor(order)==1) %>%
  select(Reading:Assignment) %>% 
  pivot_longer(everything(), names_to = 'Materials') %>%
  mutate(Item = gsub("\r\n", "@", value), 
         Materials = gsub("Collab ", "", Materials)) %>% 
  select(-value) %>%
  na.omit() %>%
  separate_rows(Item, sep = "@") %>% 
  mutate(Materials = factor(Materials, levels = c("Notes", "Reading", "Video", "Assignment"))) %>% 
  arrange(Materials) %>% 
  kbl(align="ll", escape=FALSE) %>%
  kable_paper("hover", full_width=F, html_font = "Ariel",
              position='left', font_size=16) %>% 
  column_spec(1, width = '20%', border_right = TRUE) %>% 
  row_spec(0, bold = T, color = "white", background = "#654378")
```


----

# Monday

## Introductions

* To each other
* Code of conduct
* To your learning tools and platforms
    - What is HackMD?

## Data Architecture

* What is data, where can it be found, how do we use it? 
* Data Architecture collaborative notes on HackMD

----

# Wednesday 

## Spreadsheets & Metadata

* Best practices for working with data in spreadsheets
* What is metadata and why should we care? 
* Practicing data entry (hw1)

## Explore project data

* Project data you can choose from can be found [here in Google Drive](https://drive.google.com/drive/u/0/folders/1jULudBjRbHdW-uLIvmMbxRBEJJkq9crY)
* Browse through the available data sets, read the summaries and skim the codebooks. Next week you'll choose a data set and start to develop research questions to explore. 

