---
title: "Week 02: Introduction to the Class"
output:   
  html_document: 
    toc: yes
    toc_float: true
    toc_depth: 2
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr);library(readxl); library(kableExtra); library(tidyr)
library(stringr)
options(knitr.table.format = "html", knitr.kable.NA = '') 
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE) 
```


----

# Learning Path

## Where we've been
* Learning how to interact with data.
* Exploring the data available for this semester

## Where we're at
* Choosing a data set
* Learning what is involved in data processing


## Objectives for the week

* Write down several research questions
* Import your chosen data into your SPC
* Create a plan to handle data management in a reproducible manner

---

```{r}
read_excel("schedule_615.xlsx", sheet="weekly_schedule") %>% 
  mutate(Date = format(date, "%m/%d")) %>% 
  filter(floor(order)==2) %>%
  select(Reading:Assignment) %>% 
  pivot_longer(everything(), names_to = 'Materials') %>%
  mutate(Item = gsub("\r\n", "-", value)) %>% 
  select(-value) %>%
  na.omit() %>%
  separate_rows(Item, sep = "-") %>% 
  mutate(Materials = factor(Materials, levels = c("Lecture Notes", "Collaborative Notes", "Reading", "Video", "Assignment"))) %>% 
  arrange(Materials) %>% 
  kbl(align="ll", escape=FALSE) %>%
  kable_paper("hover", full_width=F, html_font = "Ariel",
              position='left', font_size=16) %>% 
  column_spec(1, width = '20%', border_right = TRUE) %>% 
  row_spec(0, bold = T, color = "white", background = "#654378")
```


----

# Monday

## Formulating testable hypothesis

* Asking questions is easy. Asking _answerable_ questions is more difficult. 
* We'll be working in lec02: Formulating research questions lecture notes, and the corresponding HackMD collab notes

----

# Wednesday 

## Preparing data for analysis

* Do not underestimate the importance of this step
* It will take you far, far longer than you anticipate to 'clean' your data
* Writing code will be challenging, but will pay off in the long run

