---
title: "Week 03: Data Management"
output:   
  html_document: 
    toc: yes
    toc_float: true
    toc_depth: 2
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr);library(readxl); library(kableExtra); library(tidyr)
library(stringr)
options(knitr.table.format = "html", knitr.kable.NA = '') 
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE) 
```


----

# Learning Path

## Where we've been
* Choosing a data set & writing a research question
* Learning what is involved in data processing

## Where we're at
* Data wrangling

## Objectives for the week

* Create a reproducible script file that does the following: 
    - Imports the raw data
    - performs data management tasks
    - exports an analysis ready data set
    
---

```{r}
read_excel("schedule_615.xlsx", sheet="weekly_schedule") %>% 
  mutate(Date = format(date, "%m/%d")) %>% 
  filter(floor(order)==3) %>%
  select(Reading:Assignment) %>% 
  pivot_longer(everything(), names_to = 'Materials') %>%
  mutate(Item = gsub("\r\n", "@", value), 
         Materials = gsub("Collab ", "", Materials)) %>% 
  select(-value) %>%
  na.omit() %>%
  separate_rows(Item, sep = "@") %>% 
  mutate(Materials = factor(Materials, levels = c("Notes", "Reading", "Video", "Assignment"))) %>% 
  arrange(Materials) %>% 
  kbl(align="ll", escape=FALSE) %>%
  kable_paper("hover", full_width=F, html_font = "Ariel",
              position='left', font_size=16) %>% 
  column_spec(1, width = '20%', border_right = TRUE) %>% 
  row_spec(0, bold = T, color = "white", background = "#654378")
```


----

# Monday

> No Class, Labor day. 

Enjoy the three day weekend, and consider learning something new about this holiday. (Optional reading)

* [US Department of Labor: "History of Labor Day"](https://www.dol.gov/general/laborday/history)
* [America's Black Holocaust Museum: "America's Racist History of Labor"](https://www.abhmuseum.org/watch-americas-racist-history-of-labor/)
* [Smithsonian: "Striking Union Workers Turned the First Labor Day Into a Networking Event "](https://www.smithsonianmag.com/history/striking-union-workers-turned-first-labor-day-networking-event-180964705/)
* [UAW: "Celebrating Labor day by Celebrating Labor Unions"](https://uaw.org/celbrating-labor-day-celebrating-labor-unions/)

Got more? Share in our Discord channel.

----

# Wednesday 

We will be working on Homework #3 today in class. Your goal is to have a working script file before the end of the day. 

There are no new notes for this section. Continue to use the Data Management chapter of the [Applied Stats notebook](https://norcalbiostat.github.io/AppliedStatistics_notes/data-management.html) for code and examples in R. 


