---
title: "Describing distributions of data"
output:
  html_document:
    highlight: tango
    theme: flatly
  pdf_document: default
css: ../css/customh5.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Assignment Overview

There are a variety of conventional ways to visualize data - tables, histograms, bar graphs, etc. Now that your data have been cleaned up, it is time to examine the distribution of variables related to your research question. You will create a plot, follow up each graphic with a table of summary statistics (for quantitative variables) or frequency and proportion table (for categorical), and then a summary paragraph that brings it all together. 

# Instructions

**You must use your cleaned version of your research data for this!** This is the data file that was saved as a result of you running your data management script file (`dm.Rmd` or `dm.sps`). 

0. Use the template provided: [[RMD]](Univariate_Graphing_template.Rmd) for R users, and [[Word]](Univariate_Graphing_template.docx) for SPSS users. 
1. Completely describe 2 categorical and 2 quantitative variables using 
    - A table of summary statistics,
    - An appropriate plot with titles and axes labels, 
    - A short paragraph description in full complete English sentences. 
   

To guide your description of this distribution try to include the following information:

* What is the trend in the data? What exactly does the chart show? (Use the chart title to help you answer this question)
* What are the axes and what are the units?
* Describe the shape:
    - Symmetry/Skewness - Is it symmetric, skewed right, or skewed left?
    - Modality - Is it uniform, unimodal, or bimodal?             
* Describe the spread:
    - Variability - What is the approximate range of the data (x-axis)?
    - Does the variable have a lot of variability in the data (visually, are the participants responded to many different responses or mainly just one)?
* Describe the center: What is the mean/median/midpoint of the data? (Pick one or two). Don't 
* Describe the outliers (note: there may not be any for every graph):
    - Are there any outliers for the variable?
    - If yes, are these true outliers or false (due to data management or input error) outliers?

# Example
This example uses the `mpg` data set from the `ggplot2` package.
```{r}
mpg <- ggplot2::mpg
```

## Basic categorical
Draft style plot, direct computer output showing/copied. Poor grammar and/or sentence structure, no attempt at explaining what the variable means, extra unnecessary or incorrect information included. Typos. 

> class

```{r}
library(descr)
freq(mpg$class)
```

> theres more suvs than compacts. 2% are 2seaters. there are 5 2seaters 47 cmpact 41 midize 11 minivans 33 pickups 35% subcompacts, 62 suv and 234 total cars. 

## Proficient categorical
Cleaned up plot, full English sentences, useful text formatting of variable names and levels. Explained what the variable was named and what it measured. 

> The `class` variable from the `mpg` data set is a catgorical variable that describes the type of vehicle being measured. Some levels of this categorical variable include _compact_, _pickup_ and _suv_. 

```{r, fig.height=4, fig.align='center'}
library(sjPlot); library(ggplot2)
set_theme(base = theme_classic())
sjp.frq(mpg$class)
```

> Sub compact cars are the most frequently reported type of car, making up over one-quarter (26.5%) of the cars in this data set with n=62 cars represented. The least represented car is a compact car with n=5 (2.1%) records. 

## Basic quantitative 

```{r}
hist(mpg$cty)
summary(mpg$cty)
```

## Proficient quantitative
Overlaid a density curve on the histogram, also looked at a boxplot for outliers. Table of summary statistics present in a nicely formatted way, digits rounded appropriately. Plot cleaned up with appropriate axis and titles.  

> The `cty` variable records the miles per gallon (mpg) achieved during city driving. This is a quantititative numeric variable. 

```{r, fig.height=3, fig.width=4, fig.align='center'}
ggplot(mpg, aes(x=cty)) + geom_histogram(aes(y=..density..), fill="grey", binwidth = 2) + 
      geom_density()+ xlab("MPG") + ggtitle("City miles per gallon (MPG)") 
boxplot(mpg$cty)
knitr::kable(t(c(summary(mpg$cty), sd=sd(mpg$cty))), digits=1)
```

> The MPG in the city ranges from 9 to 35, unimodal and is slightly skewed right with a mean of 16.9 close to the median of 17 and a standard deviation of 4.3mpg. The boxplot indicates that there are at least 4 upper end outliers achieving a city MPG of approximately over 28 mpg. 



##### Submission
* Upload the final PDF to `hw02 Univariate Graphics/Incoming` folder in Google Drive with the file name: _userid_univ_graph.pdf_
* This assignment will be peer reviewed. 
