---
title: "Describing distributions of data"
output:
  pdf_document: default
  html_document:
    highlight: tango
    theme: flatly
    toc: true
    toc_float: true
css: ../css/custom_css.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Assignment Overview

There are a variety of conventional ways to visualize data - tables, histograms, bar graphs, etc. The purpose is always to examine the distribution of variables related to your research question. You will create a plot, follow up each graphic with a table of summary statistics (for quantitative variables) or frequency and proportion table (for categorical), and then a summary paragraph that brings it all together. 

# Instructions

* Use the template provided: [[RMD]](04_univariate_graphing_template.Rmd) for R users, and [[Word]](04_univariate_graphing_template.docx) for SPSS users. Rename this file to `univ_graphing_userid`
* Completely describe 2 categorical and 2 quantitative variables using 
    - A table of summary statistics,
    - An appropriate plot with titles and axes labels, 
    - A short paragraph description in full complete English sentences. 
* Upload your final PDF to `04 Univariate Graphing` folder in Google Drive.
* This is a peer reviewed assignment. 
* Grading rubric is available for viewing in Blackboard Learn. 

To guide your description of this distribution try to include the following information:

* What is the trend in the data? What exactly does the chart show? (Use the chart title to help you answer this question)
* Describe the location of the bulk of the data. Measures include
    - Measures of center: mean/median
    - IQR as Q3 and Q1 (i.e., 50% of the data lie between $x_{1}$ and $x_{2}$)
    - N and (%) for categorical data for the largest category. 
* Describe the shape:
    - Symmetry/Skewness - Is it symmetric, skewed right, or skewed left?
    - Modality - Is it uniform, unimodal, or bimodal?       
* Describe the spread or variability in the data
    - Appropriate measures include range, standard deviation, IQR for continuous data
    - For categorical data describe if distribution is spread across multiple response categories or mainly only one.
* Describe the outliers (note: there may not be any for every graph). Continuous data only.
    - Are there any outliers for the variable?
    - If yes, are these true outliers or false (due to data management or input error) outliers? 
    (This could alert you to missing codes like -77 or 99 that need to be set to missing)

\newpage 

# Example
This example uses the `mpg` data set from the `ggplot2` package.
```{r}
mpg <- ggplot2::mpg
```

## Basic categorical
Draft style plot, direct computer output showing/copied. Poor grammar and/or sentence structure, no attempt at explaining what the variable means, extra unnecessary or incorrect information included. Typos. 

> class

```{r}
library(descr)
freq(mpg$class)
```

> theres more suvs than compacts. 2% are 2seaters. there are 5 2seaters 47 cmpact 41 midize 11 minivans 33 pickups 35% subcompacts, 62 suv and 234 total cars. 

## Proficient categorical
Cleaned up plot, full English sentences, useful text formatting of variable names and levels. Explained what the variable was named and what it measured. 

> The `class` variable from the `mpg` data set is a catgorical variable that describes the type of vehicle being measured. Some levels of this categorical variable include _compact_, _pickup_ and _suv_. 

```{r, fig.height=4, fig.align='center'}
library(sjPlot); library(ggplot2)
set_theme(base = theme_classic())
sjp.frq(mpg$class)
```

> Sub compact cars are the most frequently reported type of car, making up over one-quarter (26.5%) of the cars in this data set with n=62 cars represented. The least represented car is a compact car with n=5 (2.1%) records. 

\newpage

## Basic quantitative 

```{r}
hist(mpg$cty)
summary(mpg$cty)
```

## Proficient quantitative
Overlaid a density curve on the histogram, also looked at a boxplot for outliers. Table of summary statistics present in a nicely formatted way, digits rounded appropriately. Plot cleaned up with appropriate axis and titles.  

> The `cty` variable records the miles per gallon (mpg) achieved during city driving. This is a quantititative numeric variable. 

```{r, fig.height=3, fig.width=4, fig.align='center'}
ggplot(mpg, aes(x=cty)) + geom_histogram(aes(y=..density..), fill="grey", binwidth = 2) + 
      geom_density()+ xlab("MPG") + ggtitle("City miles per gallon (MPG)") 
boxplot(mpg$cty)
knitr::kable(t(c(summary(mpg$cty), sd=sd(mpg$cty))), digits=1)
```

> The MPG in the city ranges from 9 to 35, unimodal and is slightly skewed right with a mean of 16.9 close to the median of 17 and a standard deviation of 4.3mpg. The boxplot indicates that there are at least 4 upper end outliers achieving a city MPG of approximately over 28 mpg. 


