---
title: "Univariate Graphing Assignment"
author: "your name here"
output:
  html_document:
    highlight: tango
    theme: flatly
  pdf_document: default
  word_document: default
---


```{r}
# put the code to read in your cleaned data here. Delete this text afterward.
```

# Name of quantitative variable 1 here
> Describe this variable in words (what does it say in the codebook) then delete this text. 

```{r}
# put your summary table code here
```

```{r}
# put your plotting code here
```

> Describe the distribution of the variable here in paragraph form. Then delete this text. 


# Name of quantitative variable 2 here
> Describe this variable in words (what does it say in the codebook) then delete this text. 

```{r}
# put your summary table code here
```

```{r}
# put your plotting code here
```

> Describe the distribution of the variable here in paragraph form. Then delete this text. 


# Name of categorical variable 1 here
> Describe this variable in words (what does it say in the codebook) then delete this text. 

```{r}
# put your frequency and proportion table code here
```

```{r}
# put your plotting code here
```

> Describe the distribution of the variable here in paragraph form. Then delete this text. 


# Name of categorical variable 2 here
> Describe this variable in words (what does it say in the codebook) then delete this text. 

```{r}
# put your frequency and proportion table code here
```

```{r}
# put your plotting code here
```

> Describe the distribution of the variable here in paragraph form. Then delete this text. 
